% On-board computer for PC104 format CubeSats
% ---------------------------------------------------------------------
% Filip Geib Apr. 2021

% export in Long-term Archiving (PDF/A)
% conversion needed: https://pdfrecover.herokuapp.com/pdfaconvert/
\input glyphtounicode.tex
\input glyphtounicode-cmr.tex
\input glyphtounicode-ntx.tex
\pdfgentounicode=1

% define English quotes
\enquotes

% include template
\input ctustyle3

\worktype [B/EN]

\faculty    {F3}
\department {Department of Measurement}
\title      {On-board computer for PC104 format CubeSats}
\subtitle   {}

\author     {Filip  Geib}
\date       {May 2021}
\supervisor {Ing. Vojtěch Petrucha, Ph.D.}
\studyinfo  {Cybernetics and Robotics}
\workname   {}

\workinfo   {} %\url{https://github.com/visionspacetec/VST104}
\titleSK    {Palubný počítač pre CubeSaty formátu PC104}
\subtitleSK {}

\pagetwo    {}



%%%%% ABSTRACT - skontrolovane

\abstractEN {
    This thesis is focused on the development of PC/104 format electronic boards for CubeSat applications. Particular attention was given to a single on-board computer (OBC) module. This universal ARM-based OBC is driven by an STM32L4 microcontroller supporting a wide range of interfaces. Its additional features include robust power management, separate peripheral isolation, triple-redundant FLASH and F-RAM memories, two CAN bus transceivers, built-in temperature monitoring, and a significant payload sector. A dedicated radiation testing was conducted under a gamma radiation source. Moreover, three additional boards were developed, including a double redundant version of the OBC, a universal PC/104 module, and a FlatSat test bench. All of these boards were designed with open-source principles in a KiCad environment. This thesis contributes to the VisionSpace Technologies VST104 project by introducing a hardware platform for mission control systems testing and compression algorithms development.
    \bigskip
}
\abstractSK {
    Táto práca je zameraná na vývoj elektronických dosiek formátu PC/104 pre CubeSat aplikácie. Osobitná pozornosť bola venovaná modulu palubného počítača (OBC). Tento univerzálny OBC založený na ARM technológii je riadený mikrokontrolérom STM32L4 podporujúcim širokú škálu rozhraní. Medzi jeho ďalšie funkcie patrí robustná správa napájania, samostatná izolácia periférii, trojito redundantné pamäte FLASH a F-RAM, dva komunikátory zbernice CAN, zabudované monitorovanie teploty a rozsiahly sektor užitočného nákladu. Cielené testovanie odolnosti voči žiareniu bolo uskutočnené pod zdrojom gama radiácie. Ďalej boli vyvinuté tri dosky, vrátane dvojito redundantnej verzie OBC, univerzálneho modulu PC/104 a testovacej platformy FlatSat. Všetky tieto dosky boli navrhnuté podľa princípov open-source v prostredí KiCad. Táto práca prispieva k projektu VisionSpace Technologies VST104 zavedením hardvérovej platformy určenej pre testovanie systémov riadenia misií a vývoja kompresných algoritmov.
    \bigskip
}



%%%%% KEYWORDS

\keywordsEN {
    CubeSat; FlatSat; OBC; PC104; radiation testing.
}
\keywordsSK {
    CubeSat; FlatSat; OBC; PC104; radiačné testovanie.
}



%%%%% ACKNOWLEDGEMENT

\thanks {
    I would like to express my gratitude to VisionSpace Technologies, namely to José Feiteirinha, for conducting and sponsoring this project. I also thank my supervisor Vojtěch Petrucha for his guidance and support.
}



%%%%% DECLARATION

\declaration {
    I hereby declare that the presented work was developed independently and that I have listed all sources of information used within it in accordance with the methodical instructions for observing the ethical principles in the preparation of university theses. 
    \bigskip
    In Prague on May 21, 2021
    \signature % makes dots
}

%%%%% MACROS

%\draft     % Uncomment this if the version of your document is working only.
%\linespacing=1.7  % uncomment this if you need more spaces between lines
                   % Warning: this works only when \draft is activated!
%\savetoner        % Turns off the lightBlue backround of tables and
                   % verbatims, only for \draft version.
%\blackwhite       % Use this if you need really Black+White thesis.
%\onesideprinting  % Use this if you really don't use duplex printing. 
\input database/glosdata % Include glossary

%%%%% MAIN TEXT

% make title page, acknowledgment, contents etc.
\makefront

% introduction chapter
\input{chapters/chap_intro} % skontrolovane

% main body
\input{chapters/chap_project} % skontrolovane
\input{chapters/chap_sierraDes} % skontrolovane
\input{chapters/chap_sierraSub} % skontrolovane
\input{chapters/chap_delta} % skontrolovane
\input{chapters/chap_foxtrot} % skontrolovane
\input{chapters/chap_sierraTest} % skontrolovane

% conclusion chapters
\input{chapters/chap_conclusion} % skontrolovane

% bibliography
\bibchap
\usebib/c (iso690) database/bibliography

% appendencies
\input{chapters/appendix}


\bye
