
%%%%% THESIS ASSIGNMENT

\app Thesis Assignment

\midinsert
    \clabel[app:thesisAssignment]{Thesis assignment}
    \picw=0.9\hsize \cinspic database/thesisAssignment.pdf
    \caption/f Assignment of this bachelor's thesis.
\endinsert



%%%%% GLOSSARY

\app Glossary
\makeglos



%%%%% VST104 PINOUT

\app VST104 pinout

\midinsert
    \clabel[app_vst:vst104_pinout]{VST104 pinout}
    \picw=0.95\hsize \cinspic figures/diagrams/vst104_pinout.pdf
    \caption/f VST104 pinout: assignment of PC/104 header pins used in the VST104 project. Legend: red - mandatory, orange - optional, green - user defined, blue - legacy pins.
\endinsert

\midinsert \clabel[tab:vst104_pinout]{VST104 pinout}
    \ctable{cl}{
        Signal name &   Signal purpouse \crl
        {\tt UART_RCS_} & \glref{UART} used explicitly for the \glref{RCS}$^*$\cr
        {\tt GLO_SYNC}  & clock signal of global synchronization \cr
        {\tt GLO_FAULT} & signal setting a global fault flag \cr
        {\tt GLO_KS_}   & standard kill switch signal (active high) \cr
        {\tt CPU_WD_}   & watchdog signal for each of the \glref{OBC}s (\glref{CPU}s) \cr
        {\tt CPU_MODE}  & \glref{OBC} selection signal used in Board Delta \cr
    }
    \caption/t A brief explanation of selected signals from the VST104 pinout. A more specific definition will be proved ed by the \glref{VST} in the future. $*$radio communication subsystem
\endinsert



%%%%% PHOTO DOCUMENTATION

\app Photo documentation

\midinsert
    \clabel[app_photo:vst104_pinout]{VST104 pinout}
    \picw=1\hsize \cinspic figures/photos/sierra_angle.pdf
    \caption/f 
\endinsert

\midinsert
    \clabel[app_photo:vst104_pinout]{VST104 pinout}
    \picw=90.17mm \cinspic figures/photos/sierra_top.pdf
    \caption/f 
\endinsert

\midinsert
    \clabel[app_photo:vst104_pinout]{VST104 pinout}
    \picw=90.17mm\cinspic figures/photos/sierra_bottom.pdf
    \caption/f 
\endinsert

\midinsert
    \clabel[app_photo:vst104_pinout]{VST104 pinout}
    \picw=1\hsize \cinspic figures/photos/sierra_macro.pdf
    \caption/f 
\endinsert

\midinsert
    \clabel[app_photo:vst104_pinout]{VST104 pinout}
    \picw=14.5cm \cinspic figures/photos/foxtrot.pdf
    \caption/f 
\endinsert


%%%%% OVERSIZED FIGURES

\app Over-sized figures

\midinsert
    \clabel[app_vis:zeroTop]{Board Zero render top}
    \picw=90.17mm \cinspic figures/model/zero_top.pdf
    \caption/f Visualization of the Board Zero captured from its top side. The dimensions of this 3D render match the actual size of the board 1:1.
\endinsert

\midinsert
    \clabel[app_vis:sierraTop]{Board Sierra render top}
    \picw=90.17mm \cinspic figures/model/sierra_top.pdf
    \caption/f Visualization of the Board Sierra captured from its top side. The dimensions of this 3D render match the actual size of the board 1:1.
\endinsert

\midinsert
    \clabel[app_vis:sierraBottom]{Board Sierra render bottom}
    \picw=90.17mm \cinspic figures/model/sierra_bottom.pdf
    \caption/f Visualization of the Board Sierra captured from its bottom side. The dimensions of this 3D render match the actual size of the board 1:1.
\endinsert

\midinsert
    \clabel[app_sch:microcontroller]{Microcontroller schematic}
    \picw=0.88\hsize \cinspic figures/schemes/microcontroller.pdf
    \caption/f Schematic diagram of the microcontroller and its auxiliaries.
\endinsert

\midinsert
    \clabel[app_vis:foxtrotTop]{Element Foxtrot render}
    \picw=14.5cm \cinspic figures/model/foxtrot_top.pdf
    \caption/f Visualization of the Element Foxtrot captured from its top side. The dimensions of this 3D render match the actual size of the board 1:1.4.
\endinsert

\midinsert
    \clabel[app_vis:foxtrotKicad]{Element Foxtrot KiCad}
    \picw=14.5cm \cinspic figures/pcbs/foxtrot_kicad.pdf
    \caption/f PCB design of the Element Foxtrot captured directly in the KiCad environment. The dimensions of this render match the actual size of the board 1:1.4.
\endinsert



%%%%% ADDITIONAL MATERIALS

\app Additional materials

\midinsert \clabel[tab:componentsConsumption]{Components consumption}
    \ctable{ccccc}{
        \vspan2{Designator} & \vspan2{Power bus  $\rm{[V]}$} & \mspan3[c]{Current consumption $\rm{[mA]}$} \cr
                    &       & Min.  & Typ.  & Max. \crl
        AS[1-15]    & 5     & -     & 0.02  & 0.06 \cr %DGQ2788A  
        EF[1,2]     & 3.3, 5 & 0.14 & 0.21  & 0.30 \cr %TPS25940-Q1
        LG1         & 3.3   & -     & 0.1   & 4    \cr %74LVC1G11-Q100
        M[1-3]      & 3.3   & 10    & 25    & 40   \cr %S25FL256L
        M[4-6]      & 3.3   & -     & 5     & 5    \cr %CY15B102Q
        TS[1-7]     & 3.3   & -     & 0.20  & 0.40 \cr %MCP9804
        U[2,3]      & 5     & -     & 40    & 70   \cr %TCAN1051V-Q1
        Y2          & 3.3   & -     & 4.0   & 4.8      %SiT8924B
    }
    \caption/t Listing of electronic components power consumption. The values were obtained from the datasheets, and correspond to the normal operation at room temperature.
\endinsert

\midinsert \clabel[tab:uncertifiedParts]{Uncertified parts variants}
    \ctable{clll}{
        Designator      & Certified part no.      & Uncertified part no.  & Difference \crl
        LG1             & 74LVC1G11GW-Q100        & 74LVC1G11GW           & auto. \crl
        \vspan2{M[1-3]} & \vspan2{S25FL256LAGNFN} & S25FL256LAGNFI        & temp. \cr
                        &                         & S25FL256LDPNFI        & temp., speed \crl
        TS[1-7]         & MCP9804x-E/MC           & MCP9808x-E/MC         & accuracy \crl
        \vspan2{Q[1-3]} & \vspan2{TPS22965W-Q1 }  & TPS22965-Q1           & temp. \cr
                        &                         & TPS22975              & temp., auto.\cr
    }
    \caption/t List of available uncertified variants to some of the OBCs electronic components. Legend: auto. - missing \glref{AEC} certification, temp. - shrink operational temperature range, speed - decreased frequency, accuracy - decreased accuracy of measurements.
\endinsert