
\chap Project description
\sec CubeSat concept
\sec PCI104 standard
\secc Mechanical specification
\secc Main header pinout
\sec OBC requirements

%%%%% FEATURES

\secc Capabilities and features
\label[chap3:features]

\begitems
    * {\sbf External clock sources:} The \glref{MCU} has two internal resistor-capacitor (\glref{RC}) oscillators that can be used to drive a master system clock and auxiliary clocks \cite[dat:mcu]. These internal oscillators have a significantly lower frequency stability and a higher temperature dependency than the external ones \cite[app:oscInt]. Therefore, to ensure the clock reliability in the harsh space environment, we had to implement external clock sources.
    * {\sbf Robust power management:} The electric power subsystem (\glref{EPS}) is known to be the most vital subsystem of a spacecraft. Its reliability and error handling should be ensured by the power control and distribution unit (\glref{PDCU}). However, it is a good engineering practice by professional \glref{OBC} manufacturers to include an additional power control to their designs \cite[obc:dataPatterns, obc:isis, obc:mit, obc:pumpkin, obc:gauss, obc:aac, obc:anteleope, obc:nanosatpro]. Our \glref{OBC} requires $3.3\rm{[V]}$ and $5\rm{[V]}$ power inputs from the main power buses. In the case of their malfunction, it is our responsibility to sense it and power down the \glref{OBC}. This feature is also important for some of the \glref{OBC} on-earth user cases. During a hardware development or a system presentation, the user might misconnect the power line or use an unsupported power source.
    * {\sbf Isolation of the peripherals:} The spacecraft \glref{OBC} is connected to many data buses shared among all other subsystems. In some scenarios, the \glref{OBC} must be able to isolate itself from a specific or multiple data buses. For example, to switch between \glref{OBC}s in a redundant configuration, handle the failure on a data bus, or prevent unintentional interferences. Standard approaches to addressing this isolation feature are analog switches \cite[the:wurzburg], optocouplers \cite[pap:satelliteLeo], or \glref{FPGA}s \cite[obc:dataPatterns]. These isolators should also guarantee that all data lines are high impedance when the \glref{OBC} is powered off.
\enditems

\secc Components certification